Space is the Place: Effects of Continuous Spatial Structure on Analysis of Population Genetic Data

### 2020-01-31 ms/second-review-responses.tex

%%%%%%
%%
%%  Don't reorder the reviewer points; that'll mess up the automatic referencing!
%%
%%%%%

\begin{minipage}[b]{2.5in}
  Resubmission Cover Letter \\
    {\it Genetics}
\end{minipage}
\hfill
\begin{minipage}[b]{2.5in}
    C.~J.~Battey, \\
    Peter Ralph, \\
    \emph{and} Andrew Kern \\
  \today
\end{minipage}

\vskip 2em

\noindent
{\bf To the Editor(s) -- }

\vskip 1em

We are writing to submit another revised version of our manuscript, 
``Space is the Place: Effects of Continuous Spatial Structure on Analysis of Population Genetic Data''.

\vspace{5em}

\noindent \hspace{4em}
\begin{minipage}{3in}
\noindent
{\bf Sincerely,}

\vskip 2em

{\bf 
C.~J.~Battey, Peter Ralph, and Andrew Kern
}\\
\end{minipage}

\vskip 4em

\pagebreak
\setcounter{page}{1}

%%%%%%%%%%%%%%
\reviewersection{AE}

The point of R3 on the bimodality in the realized dispersal distance in your model is interesting. You can make this more explicit in the paper and describe how there are realistic biological scenarios that motivate such a model (many/most plants) and demarcate how well you think you can generalize to others (the new appendix already helps in that regard). It strikes me your model has relevance for patrilocal/matrilocal systems where the paternal/maternal parent-offspring distances differ - though your model is hermaphroditic so it's an imperfect analogy. Perhaps, to explore more typical animal mating patterns, you can quickly check the impact for one or two outcomes of changing the simulation such that offspring are dispersed around the midpoint of the parental origins, as this would remove the bimodality. That said, please be sure we respect the work done is extensive and don't mean to invite/require a complete re-working. The core tension is between ensuring your results are generalizable if that is what you wish to claim; versus specific to a particular model that is biologically motivated. It's particularly key for the material on GWAS, where you are explicitly trying to address human GWAS results: those sections will be read and taken at face value by a large community even if this simulation scheme introduces some discrepancies from what might more realistically emerge. So at a minimum, some caveats directly in the GWAS discussion section would help.

%%%%%%%%%%%%%%
\reviewersection{1}

\begin{quote}
Only minor changes were needed, and they were made. This study neatly summarises many biases arising from unmodelled continuous space (pertinent for real studies), and I would not extend it by further analyses. When polishing the final version, the authors can make some of their messages stronger. For example, Figure 7 and the accompanying text may give the impression that we are doing pretty well with smc++. It is only when panel B is seen one can appreciate the degree of uncertainty. So, in a way, Figure S7 is more telling as to how much we can trust our results in real analysis. After all, in reality, as authors mention, we only have one sample set, not replicates.
\end{quote}

Thanks!!!

%%%%%%%%%%%%%%
\reviewersection{2}

\begin{quote}
The authors have done a commendable job of responding to the reviewers' comments. The extended discussion and additional supplemental figures helps with the clarity of the manuscript. The results for the comparison with the discrete stepping-stone model were useful - though see below. The authors have also provided sufficient detail justifying their choice of model parameters.
\end{quote}

\begin{quote}
I don't think I have any critical remaining issues or suggestions, but I did have a couple of thoughts. The authors might just want to think about them.
\end{quote}

\begin{point}{}
I appreciate the appendix comparing summary statistics in the continuous and stepping-stone models, although I must confess I am now even more uncertain than I was before about which models are more appropriate. I get that the coalescent models make assumptions that are not directly interpretable in terms of observable parameters (although are "interactions between geographically distant parts of the species range" really biologically unrealistic?). This forward model is more interpretable, in that sense, but does it actually produce more realistic data? I guess that's a question for future work, but that question was one of the main things I came away from this paper with.
\end{point}

\reply{
}

\begin{point}{\revref}
 "PCA is unable to capture". Is this a sample size effect? i.e. the precision of the estimates of the PCs depends on sample size, so it's possible that you just need a larger sample to estimate the PCs when you have weak structure. Or is it something else?
\end{point}

\reply{
}

\begin{point}{Line 685.}
 Is Khera et al (2018) really the right reference for the idea of polygenic scores? They have been around for a lot longer than that.
\end{point}

\reply{
}

\begin{point}{Line 686. }
 I don't know that the results of Berg and Sohail et al. are indicating a fundamental issue with PC correction, rather than a problem with meta-analysis.
\end{point}

\reply{
}

\begin{point}{Line 689-92. }
 Is the result of Martin et al. really due to population stratification? There seem to be a number of factors affecting transferability of PRS (LD differences, effect size heterogeneity, differences in causal variants etc..), but I don't know that anyone thinks residual structure is the major one, particularly for eg. UKB summary stats.
\end{point}

\reply{
}

\begin{point}{Line 752 "population genetics flavor"}
I don't understand what this means?
\end{point}

\reply{
}

\begin{point}{Line 800}
"process-driven descriptions of ancestry and/or more generalized unsupervised methods". Aren't these two things opposite? That's ok, but it sounds a bit like you have no idea what would help...
\end{point}

\reply{
}

\begin{point}{}
 Fix citation: Peter L. Ralph and Ashander ??? on pg. 5, line 216
\end{point}

\reply{
}

\begin{point}{}
'there' written twice on pg. 27, line 1033
\end{point}

\reply{
}

%%%%%%%%%%%%%%
\reviewersection{3}

\begin{quote}
I am fairly pleased overall with the revised version of this article. I commend the authors for the substantial amount of work conducted. This contribution now convincingly demonstrates how important it is to properly account for population structure when individuals are distributed along a spatial continuum.
\end{quote}

\begin{point}{}
A couple of minor points about the forward model are still puzzling me though. First of all, in one dimension, the variance of $L(o)$, the random variable corresponding to the offspring's spatial position, given $l(p)$, the parent's position, is equal to $E((L(o)-l(p))^2):=\sigma^2$. Hence, $\sigma$ is here the square root of the expected squared Euclidean distance between parents and offspring. In general, assuming isotropic migrations, $\sigma$ is the square root of the expected squared Euclidean distance taken along any dimension. It seems to me that this definition is distinct from the mean parent-offspring distance, as stated by the authors.
\end{point}

\reply{
}

\begin{point}{}
More importantly, I am still a bit confused about the reproduction scheme. My understanding is that, once mating takes place, the two parents involved (noted as P(1) and P(2)) each produce a Poisson number of offspring. The location of each offspring is determined by a Gaussian distribution of covariance matrix $\sigma^2 I$, centered on the location of the parent that ``produced' this offspring. Let $O(1,i)$ and $O(2,j)$ be the i-th and j-th offspring produced by $P(1)$ and $P(2)$ respectively. Now, if I am not mistaken, half of the genetic material of O(2,j) comes from P(1) (the other half coming from P(2)) while O(1,i) receives one chromosome from P(2) (and another one from its ``true' parent, i.e., P(1)). Hence, when it comes to the parent-offspring distance as measured from their genetic material, half of the chromosomes have a ``small' dispersal distance, equal to that of the parent-offspring distance, while the other half have a ``large' distance roughly equal to the parent-offspring distance plus the distance between mates.
\end{point}

\reply{
}

\begin{point}{}
It is important that the authors clarify this last point because most results presented in this study rely on Wright's neighborhood size, which derives from a model where dispersal of individuals and chromosomes follow well-defined spatial dynamics. A sensible comparison of neighborhood sizes requires that dispersal (along with effective population density) is quantified in a coherent way across different models.
\end{point}

\reply{
}

\begin{point}{}
Additionally, the neighborhood size as defined by the authors (see line 211) is a function of the census density rather than the effective density of individuals. I am therefore not fully convinced that the range of values for this parameter is comparable to that given in Table 1. Furthermore, the neighborhood size estimates obtained for Bebicium vittatum derive from the analysis of a one-dimensional habitat. The product $\sigma^2 \rho$ is thus expressed in this particular case as a number of individuals *per unit of space* (in two dimensions, this product is expressed as a number of individuals). Hence, the unit of space matters a lot here, prohibiting the direct comparison with other figures in the same table or with the simulations conducted in this study.
\end{point}

\reply{
}

\begin{point}{}
Finally, Leblois, Estoup and Streiff (2006) Molecular Ecology conducted a large simulation studies with a focus very close to that of the present study. In particular, they investigated the impact of sampling on the inference of summary statistics commonly used in population genetics. It would probably be relevant to compare some of the results presented in Leblois et al. to that put forward here.
\end{point}

\reply{
}

%%%%%%%%%%%%%%
\reviewersection{4}

\begin{quote}
The authors have thoroughly addressed nearly all of my comments in their timely manuscript, and I am satisfied with their responses.
\end{quote}

\begin{point}{}
Regarding Table S1, I would still suggest adding a column about interpretation for those variables and evaluated with respect to the simulated results.
\end{point}

\reply{
}

### 2020-02-10 ms/spaceness.tex
draft response 2

% The best-studied approaches to population genetics in continuous space were developed by \citet{Wright1942,Wright1943} and \citet{Malecot1948}, who derived expressions for genetic differentiation in continuous space assuming Poisson distributed numbers of offspring and independent dispersal among individuals. 
The degree to which genetic relationships are geographically correlated
depends on the chance that two geographically nearby individuals are close relatives
-- in modern terms, by the tension between
migration (the chance that one is descended from a distant location)
and coalescence (the chance that they share a parent).
A key early observation by \citet{Wright1946} is that this balance is often nicely summarized by the ``neighborhood size'',
defined to be $N_W = 4 \pi \rho \sigma^2$, where $\sigma$ is the mean parent--offspring distance and $\rho$ is population density. 
This can be thought of as proportional to the average number of potential mates for an individual (those within distance $2 \sigma)$,
or the number of potential parents of a randomly chosen individual. Empirical estimates of neighborhood size vary hugely across species
-- even in human populations,
estimates range from 40 to over 5,000 depending on the population and method of estimation (Table~\ref{table:NStable}).
%\citet{Maruyama1972} found that the rate of decline in genetic diversity in a two-dimensional continuous population approaches the random mating expectation when $\rho\sigma^2 > 1$, and proposed that this had the important implication that most population genetic expectations for randomly mating populations could be applied to continuously distributed populations with relatively little error.

@@

We simulated populations using the program SLiM v3.1 \citep{Haller2019}. 
Each time step consists of three stages: reproduction, dispersal, and mortality. 
To reduce the parameter space we use the same parameter, denoted $\sigma$, 
to modulate the spatial scale of interactions at all three stages by adjusting the standard deviation of the corresponding Gaussian functions. 
As in previous work \citep{Wright1943,Ringbauer2017}, $\sigma$ is equal to the mean parent-offspring distance.

@@

Clinal environments cause an interesting pattern in false positives after PC correction: at low neighborhood sizes the correction removes nearly all significant associations, 
but at neighborhood sizes above roughly 250 the proportion of significant SNPs increases to up to 0.4\% (Figure~\ref{fig:gwas}). 
This may be due to a loss of descriptive power of the PCs
-- as neighborhood size increases, the total proportion of variance explained by the first 10 PC axes declines from roughly 10\% to 4\% (Figure \ref{fig:gwas}B). 
Essentially, PCA seems unable to effectively summarize the weak population structure present in large-neighborhood simulations, but these populations continue to have enough spatial structure to create significant correlations between genotypes and the environment. A similar process can also be seen in the corner phenotype distribution, in which the count of significant SNPs initially declines as neighborhood size increases and then increases at approximately the point at which the proportion of variance explained by PCA approaches its minimum.

@@

Previous studies have found that population structure and nonrandom sampling can create spurious signals of population bottlenecks when attempting to infer demographic history with microsatellite variation, summary statistics, or runs of homozygosity \citep{Chikhi2010,Stadler2009,Ptak2002,Mazet2015}. Here we found that methods that infer detailed population trajectories through time based on the SFS and patterns of LD across the genome are also subject to this bias, with some combinations of dispersal and sampling strategy systematically inferring deep recent population bottlenecks and overestimating ancient $N_{e}$ by around a factor of 2. We were surprised to see that both stairwayplot and SMC++ can tolerate relatively strong isolation by distance -- i.e., neighborhood sizes of 20 -- and still perform well when averaging results across multiple simulations. Inference in populations with neighborhood sizes over 20 was relatively unbiased unless samples were concentrated in the middle of the range (Figure \ref{fig:demography}). 
Although median demography estimates across many independent simulations were fairly accurate,
empirical work has only a single estimate to work with,
and individual model fits (Figure \ref{fig:demography_supplement})
suggest that spuriously inferred population size changes and bottlenecks are common, especially at small neighborhood sizes.
As we will discuss below, most empirical estimates of neighborhood size, including all estimates for human populations, are large enough that population size trajectories inferred by these approaches should not be strongly affected by spatial biases created by dispersal in continuous landscapes. In contrast, \citet{Mazet2015} found that varying migration rates through time could create strong biases in inferred population trajectories from an $n$-island model with parameters relevant for human history, suggesting that changes in migration rates through time are more likely to drive variation in inferred $N_{e}$ than isolation by distance.

@@

In this work we simulated a single chromosome with size roughly comparable to one human chromosome. If we scale the number of false-positive associations identified in our analyses to a GWAS conducted on whole-genome data from humans, we would expect to see several thousand weak false-positive associations after PC corrections in a population with neighborhood sizes up to at least 1000 (which should include values appropriate for many human populations). Notably, very few of the spurious associations we identified would be significant at a conservative Bonferroni-adjusted $p$-value cutoff (see Figure \ref{fig:manhattan_plots}). This suggests that GWAS focused on finding strongly associated alleles for traits controlled by a limited number of variants in the genome are likely robust to the impacts of continuous spatial structure. However, methods that analyze the combined effects of thousands or millions of weakly associated variants such as polygenic risk scores \citep{Khera2018} are likely to be affected by subtle population structure. Indeed as recently identified in studies of genotype associations for human height in Europe \citep{Berg2018,Sohail2018}, PC regression GWAS in modern human populations do include residual signal of population structure in large-scale analyses of polygenic traits. When attempting to make predictions across populations with different environmental exposures, polygenic risk scores affected by population structure can be expected to offer low predictive power, as was shown in a recent study finding lower performance outside European populations \citep{Martin2019}.

@@

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Thu Apr 11 13:11:59 2019
\begin{table}[hbpt]
\small
\begin{adjustwidth}{-.9cm}{}
\centering
\caption{\bf Neighborhood size estimates from empirical studies.}
\begin{tabular}{rllrrll}
  \hline
 Species & Description & \makecell[l]{Neighborhood\\Size} & Method & Citation \\ 
  \hline
  \makecell[l]{\textit{Ipomopsis}\\\textit{aggregata}} & flowering plant & 12.60 - 37.80 & Genetic & \citep{Campbell1992}\\
  %\makecell[l]{\textit{Linanthus}\\\textit{parryae}} & flowering plant & 14 - 27 & Genetic & \citep{Wright1943}\\ %this one got debunked by a series of studies in the 90's that found that spatial differentiation in flower color is driven by environmental variability across years. See https://doi.org/10.1111/j.1558-5646.2007.00219.x
  \makecell[l]{\textit{Borrichia}\\\textit{frutescens}} & salt marsh plant & 20 - 30 & Genetic+Survey & \citep{Antlfinger1982} \\ 
  \makecell[l]{\textit{Oreamnos}\\\textit{americanus}} & mountain goat & 36 - 100 & Genetic & \citep{Shirk2014} \\ 
  \makecell[l]{\textit{Homo sapiens}} & \makecell[l]{Gainj- and Kalam- \\speaking people,\\Papua New Guinea} & 40 - 213 & Genetic & \citep{Rousset1997}\\ 
  \makecell[l]{\textit{Formica sp.}} & colonial ants & 50 - 100 & Genetic & \citep{Pamilo1983} \\ 
  \makecell[l]{\textit{Astrocaryum}\\\textit{mexicanum}} & palm tree & 102 - 895 & Genetic+survey & \citep{Eguiarte1993} \\ 
  \makecell[l]{\textit{Spermophilus}\\\textit{mollis}} & ground squirrel & 204 - 480 & Genetic+Survey & \citep{Antolin2001}\\
  \makecell[l]{\textit{Sceloperus}\\\textit{olivaceus}} & lizard & 225 - 270 & Survey & \citep{Kerster1964} \\ 
  \makecell[l]{\textit{Dieffenbachia}\\\textit{longispatha}} & \makecell[l]{beetle-pollinated\\colonial herb} & 227 - 611 & Survey & \citep{Young1988} \\ 
  \makecell[l]{\textit{Aedes aegypti}} & \makecell[l]{Yellow-fever mosquito} & 268 & Genetic & \citep{Jasper2019} \\ 
  \makecell[l]{\textit{Homo sapiens}} & \makecell[l]{Gainj- and Kalam- \\speaking people,\\Papua New Guinea} & 410 & Survey & \citep{Rousset1997} \\ 
  \makecell[l]{\textit{Quercus}\\\textit{laevis}} & Oak tree & $>440$ & Genetic & \citep{Berg1995} \\ 
  \makecell[l]{\textit{Drosophila}\\\textit{pseudoobscura}} & fruit fly & 500 - 1,000 & Survey+Crosses & \citep{Wright1946} \\ 
  \makecell[l]{\textit{Homo sapiens}} & \makecell[l]{POPRES data\\NE Europe} & 1,342 - 5,425 & Genetic & \citep{Ringbauer2017} \\ 
  \makecell[l]{\textit{Bebicium}\\\textit{vittatum}} & intertidal snail & 240,000 & Survey & \citep{Rousset1997}\\ 
  \makecell[l]{\textit{Bebicium}\\\textit{vittatum}} & intertidal snail & 360,000 & Genetic & \citep{Rousset1997}\\ 
   \hline
\end{tabular}
\label{table:NStable}
\end{adjustwidth}
\end{table}

\subsection{Where are natural populations on this spectrum?}
For how much of the tree of life do spatial patterns circumscribe genomic variation?
In Table \ref{table:NStable} we gathered estimates of neighborhood size from a range of organisms
to get an idea of how strongly local geographic dispersal affects patterns of variation.
This is an imperfect measure:
some aspects of genetic variation are most strongly determined by neighborhood size \citep{Wright1946},
others (e.g., number of segregating sites)
are more strongly determined by global $N_e$ or by the ratio of the two.
% These values should be compared to our simulation results with some caution, as though we expect neighborhood size to have similar effects across species of varying global $N_e$ \citep{Wright1946}, in our study we evaluated only a relatively small population of $\approx 10,000$. 
\llabel{popsizecaveat}
In addition, these empirical examples are likely biased towards small-neighborhood species (because few studies have quantified neighborhood size in species with very high dispersal or population density). However, from the available data we find that neighborhood sizes in the range we simulated are fairly common across a range of taxa. At the extreme low end of empirical neighborhood size estimates we see some flowering plants, large mammals, and colonial insects like ants. Species such as this have neighborhood size estimates small enough that spatial processes are likely to strongly influence inference.
These include some human populations such as the Gainj- and Kalam-speaking people of Papua New Guinea, in which the estimated neighborhood sizes in \citet{Rousset1997} range from 40 to 410 depending on the method of estimation. 
Many more species occur in a middle range of neighborhood sizes between 100 and 1000 -- a range in which spatial processes play a minor role in our analyses under random spatial sampling but are important when sampling of individuals in space is clustered. Surprisingly, even some flying insects with huge census population sizes fall in this group, including fruit flies (\textit{D.~melanogaster}) and mosquitoes (\textit{A.~aegypti}). Last, many species likely have neighborhood sizes much larger than we simulated, including the recent ancestors of modern humans in northeastern Europe \citep{Ringbauer2017}. For these species demographic inference and summary statistics are likely to reflect minimal bias from spatial effects as long as dispersal is truly continuous across the landscape. While that is so we caution that association studies in which the effects of population structure are confounded with spatial variation in the environment are still sensitive to dispersal even at these large neighborhood sizes.

@@

One facet of spatial variation that we did not address in this study is the confounding of dispersal and population density implicit in the definition of Wright's neighborhood size. Our simulations were run under constant densities, but \citet{guindon2016demographic} and \citet{Ringbauer2017} 
have shown that these parameters are identifiable under some continuous models. Similarly, though the scaling effects of dispersal we show in Figure \ref{fig:sumstats} should occur in populations of any total size, other aspects such as the number of segregating sites are also likely affected by the total landscape size (and so total census size). \llabel{landscapesizecaveat}
Much additional work remains to be done to better understand how these parameters interact to shape genetic variation in continuous space, which we leave to future studies.

>>>>

% The best-studied approaches to population genetics in continuous space were developed by \citet{Wright1942,Wright1943} and \citet{Malecot1948}, who derived expressions for genetic differentiation in continuous space assuming Poisson distributed numbers of offspring and independent dispersal among individuals. 
The degree to which genetic relationships are geographically correlated
depends on the chance that two geographically nearby individuals are close relatives
-- in modern terms, by the tension between
migration (the chance that one is descended from a distant location)
and coalescence (the chance that they share a parent).
A key early observation by \citet{Wright1946} is that this balance is often nicely summarized by the ``neighborhood size'',
defined to be $N_W = 4 \pi \rho \sigma^2$, where $\sigma$ is the mean parent--offspring distance along each of the x and y axes and $\rho$ is population density (see \citep{Rousset1997} for further discussion of parameter defiitions in one- and two-dimensional habitats) \llabel{axialdist}. 
This can be thought of as proportional to the average number of potential mates for an individual (those within distance $2 \sigma)$,
or the number of potential parents of a randomly chosen individual. Empirical estimates of neighborhood size vary hugely across species
-- even in human populations,
estimates range from 40 to over 5,000 depending on the population and method of estimation (Table~\ref{table:NStable}).
%\citet{Maruyama1972} found that the rate of decline in genetic diversity in a two-dimensional continuous population approaches the random mating expectation when $\rho\sigma^2 > 1$, and proposed that this had the important implication that most population genetic expectations for randomly mating populations could be applied to continuously distributed populations with relatively little error.

@@

We simulated populations using the program SLiM v3.1 \citep{Haller2019}. 
Each time step consists of three stages: reproduction, dispersal, and mortality. 
To reduce the parameter space we use the same parameter, denoted $\sigma$, 
to modulate the spatial scale of interactions at all three stages by adjusting the standard deviation of the corresponding Gaussian functions. 
As in previous work \citep{Wright1943,Ringbauer2017}, $\sigma$ is equal to the mean parent-offspring distance along each of the x and y axes.

@@

Clinal environments cause an interesting pattern in false positives after PC correction: at low neighborhood sizes the correction removes nearly all significant associations, 
but at neighborhood sizes above roughly 250 the proportion of significant SNPs increases to up to 0.4\% (Figure~\ref{fig:gwas}). 
This may be due to a loss of descriptive power of the PCs
-- as neighborhood size increases, the total proportion of variance explained by the first 10 PC axes declines from roughly 10\% to 4\% (Figure \ref{fig:gwas}B). 
Essentially, PCA seems unable to effectively summarize the weak population structure present in large-neighborhood simulations given the sample sizes we tested, \revpoint{2}{2} \llabel{samplesizenote}
but these populations continue to have enough spatial structure to create significant correlations between genotypes and the environment. A similar process can also be seen in the corner phenotype distribution, in which the count of significant SNPs initially declines as neighborhood size increases and then increases at approximately the point at which the proportion of variance explained by PCA approaches its minimum.

@@

Previous studies have found that population structure and nonrandom sampling can create spurious signals of population bottlenecks when attempting to infer demographic history with microsatellite variation, summary statistics, or runs of homozygosity \citep{Chikhi2010,Stadler2009,Ptak2002,Mazet2015,leblois2006genetics}. Here we found that methods that infer detailed population trajectories through time based on the SFS and patterns of LD across the genome are also subject to this bias, with some combinations of dispersal and sampling strategy systematically inferring deep recent population bottlenecks and overestimating ancient $N_{e}$ by around a factor of 2. We were surprised to see that both stairwayplot and SMC++ can tolerate relatively strong isolation by distance -- i.e., neighborhood sizes of 20 -- and still perform well when averaging results across multiple simulations. Inference in populations with neighborhood sizes over 20 was relatively unbiased unless samples were concentrated in the middle of the range (Figure \ref{fig:demography}). 
Although median demography estimates across many independent simulations were fairly accurate,
empirical work has only a single estimate to work with,
and individual model fits (Figure \ref{fig:demography_supplement})
suggest that spuriously inferred population size changes and bottlenecks are common, especially at small neighborhood sizes. 
As we will discuss below, most empirical estimates of neighborhood size, including all estimates for human populations, are large enough that population size trajectories inferred by these approaches should not be strongly affected by spatial biases created by dispersal in continuous landscapes. In contrast, \citet{Mazet2015} found that varying migration rates through time could create strong biases in inferred population trajectories from an $n$-island model with parameters relevant for human history, suggesting that changes in migration rates through time are more likely to drive variation in inferred $N_{e}$ than isolation by distance.

@@

In this work we simulated a single chromosome with size roughly comparable to one human chromosome. If we scale the number of false-positive associations identified in our analyses to a GWAS conducted on whole-genome data from humans, we would expect to see several thousand weak false-positive associations after PC corrections in a population with neighborhood sizes up to at least 1000 (which should include values appropriate for many human populations). Notably, very few of the spurious associations we identified would be significant at a conservative Bonferroni-adjusted $p$-value cutoff (see Figure \ref{fig:manhattan_plots}). This suggests that GWAS focused on finding strongly associated alleles for traits controlled by a limited number of variants in the genome are likely robust to the impacts of continuous spatial structure. However, methods that analyze the combined effects of thousands or millions of weakly associated variants such as polygenic risk scores \citep{Khera2018} are likely to be affected by subtle population structure. Indeed as recently identified in studies of genotype associations for human height in Europe \citep{Berg2018,Sohail2018}, PC regression GWAS in modern human populations do include residual signal of population structure in large-scale analyses of polygenic traits. In addition to error associated with varying patterns of linkage disequilibrium and allele frequency among populations, the confounding of environmental and genetic effects on phenotypes introduced by population structure is expected to lead to low predictive power when polygenic scores are generated for populations outside the original GWAS cohort, \llabel{ldcaveats}
%When attempting to make predictions across populations with different environmental exposures, polygenic risk scores affected by population structure can be expected to offer low predictive power, 
as was shown in a recent study finding lower polygenic score predictive power outside European populations \citep{Martin2019}.

@@

% latex table generated in R 3.5.1 by xtable 1.8-3 package
% Thu Apr 11 13:11:59 2019
\begin{table}[hbpt]
\small
\begin{adjustwidth}{-.9cm}{}
\centering
\caption{\bf Neighborhood size estimates from empirical studies.}
\begin{tabular}{rllrrll}
  \hline
 Species & Description & \makecell[l]{Neighborhood\\Size} & Method & Citation \\ 
  \hline
  \makecell[l]{\textit{Ipomopsis}\\\textit{aggregata}} & flowering plant & 12.60 - 37.80 & Genetic & \citep{Campbell1992}\\
  %\makecell[l]{\textit{Linanthus}\\\textit{parryae}} & flowering plant & 14 - 27 & Genetic & \citep{Wright1943}\\ %this one got debunked by a series of studies in the 90's that found that spatial differentiation in flower color is driven by environmental variability across years. See https://doi.org/10.1111/j.1558-5646.2007.00219.x
  \makecell[l]{\textit{Borrichia}\\\textit{frutescens}} & salt marsh plant & 20 - 30 & Genetic+Survey & \citep{Antlfinger1982} \\ 
  \makecell[l]{\textit{Oreamnos}\\\textit{americanus}} & mountain goat & 36 - 100 & Genetic & \citep{Shirk2014} \\ 
  \makecell[l]{\textit{Homo sapiens}} & \makecell[l]{Gainj- and Kalam- \\speaking people,\\Papua New Guinea} & 40 - 213 & Genetic & \citep{Rousset1997}\\ 
  \makecell[l]{\textit{Formica sp.}} & colonial ants & 50 - 100 & Genetic & \citep{Pamilo1983} \\ 
  \makecell[l]{\textit{Astrocaryum}\\\textit{mexicanum}} & palm tree & 102 - 895 & Genetic+survey & \citep{Eguiarte1993} \\ 
  \makecell[l]{\textit{Spermophilus}\\\textit{mollis}} & ground squirrel & 204 - 480 & Genetic+Survey & \citep{Antolin2001}\\
  \makecell[l]{\textit{Sceloperus}\\\textit{olivaceus}} & lizard & 225 - 270 & Survey & \citep{Kerster1964} \\ 
  \makecell[l]{\textit{Dieffenbachia}\\\textit{longispatha}} & \makecell[l]{beetle-pollinated\\colonial herb} & 227 - 611 & Survey & \citep{Young1988} \\ 
  \makecell[l]{\textit{Aedes aegypti}} & \makecell[l]{Yellow-fever mosquito} & 268 & Genetic & \citep{Jasper2019} \\ 
  \makecell[l]{\textit{Homo sapiens}} & \makecell[l]{Gainj- and Kalam- \\speaking people,\\Papua New Guinea} & 410 & Survey & \citep{Rousset1997} \\ 
  \makecell[l]{\textit{Quercus}\\\textit{laevis}} & Oak tree & $>440$ & Genetic & \citep{Berg1995} \\ 
  \makecell[l]{\textit{Drosophila}\\\textit{pseudoobscura}} & fruit fly & 500 - 1,000 & Survey+Crosses & \citep{Wright1946} \\ 
  \makecell[l]{\textit{Homo sapiens}} & \makecell[l]{POPRES data\\NE Europe} & 1,342 - 5,425 & Genetic & \citep{Ringbauer2017} \\ 
  %\makecell[l]{\textit{Bebicium}\\\textit{vittatum}} & intertidal snail & 240,000 & Survey & \citep{Rousset1997}\\ 
  %\makecell[l]{\textit{Bebicium}\\\textit{vittatum}} & intertidal snail & 360,000 & Genetic & \citep{Rousset1997}\\ 
   \hline
\end{tabular}
\label{table:NStable}
\end{adjustwidth}
\end{table}

\subsection{Where are natural populations on this spectrum?}
For how much of the tree of life do spatial patterns circumscribe genomic variation?
In Table \ref{table:NStable} we gathered estimates of neighborhood size from a range of organisms
to get an idea of how strongly local geographic dispersal affects patterns of variation.
This is an imperfect measure:
some aspects of genetic variation are most strongly determined by neighborhood size \citep{Wright1946},
others (e.g., number of segregating sites)
are more strongly determined by global $N_e$ or by the ratio of the two, and definitions of "population density" in genetic versus ecological studies may lead to varying estimates of neighborhood size for a given species \llabel{rhocaveat}. 
% These values should be compared to our simulation results with some caution, as though we expect neighborhood size to have similar effects across species of varying global $N_e$ \citep{Wright1946}, in our study we evaluated only a relatively small population of $\approx 10,000$. 
\llabel{popsizecaveat}
In addition, these empirical examples are likely biased towards small-neighborhood species (because few studies have quantified neighborhood size in species with very high dispersal or population density). However, from the available data we find that neighborhood sizes in the range we simulated are fairly common across a range of taxa. At the extreme low end of empirical neighborhood size estimates we see some flowering plants, large mammals, and colonial insects like ants with neighborhood sizes less than roughly 100. Species such as this have neighborhood size estimates small enough that spatial processes are likely to strongly influence inference.
These include some human populations such as the Gainj- and Kalam-speaking people of Papua New Guinea, in which the estimated neighborhood sizes in \citet{Rousset1997} range from 40 to 410 depending on the method of estimation. 
Many more species occur in a middle range of neighborhood sizes between 100 and 1000 -- a range in which spatial processes play a minor role in our analyses under random spatial sampling but are important when sampling of individuals in space is clustered. Surprisingly, even some flying insects with huge census population sizes fall in this group, including fruit flies (\textit{D.~melanogaster}) and mosquitoes (\textit{A.~aegypti}). Last, many species likely have neighborhood sizes much larger than we simulated, including the recent ancestors of modern humans in northeastern Europe \citep{Ringbauer2017}. For these species demographic inference and summary statistics are likely to reflect minimal bias from spatial effects as long as dispersal is truly continuous across the landscape. While that is so we caution that association studies in which the effects of population structure are confounded with spatial variation in the environment are still sensitive to dispersal even at these large neighborhood sizes.

@@

One facet of spatial variation that we did not address in this study is the confounding of dispersal and population density implicit in the definition of Wright's neighborhood size. Our simulations were run under constant densities, but \citet{guindon2016demographic} and \citet{Ringbauer2017} 
have shown that these parameters are identifiable under some continuous models. Similarly, though the scaling effects of dispersal we show in Figure \ref{fig:sumstats} should occur in populations of any total size, other aspects such as the number of segregating sites are also likely affected by the total landscape size (and so total census size). \llabel{landscapesizecaveat} 
Indeed, our finding that stepping-stone models reproduce only some of the effects on genetic variation seen in a continuous-space model (Figure \ref{fig:grid_stats}) suggests that qualitatively similar models can produce different results dependent on the specific parameterizations used. While we believe our continuous model is a more appropriate depiction of evolution in a continuous landscape than lattice models, it is likely that some populations and breeding systems do more closely resemble a series of interconnected random-mating populations. As with all population models, the best approximation for any empirical system will depend on the natural history of the species in question. \llabel{bestmodelcaveat}
Much additional work remains to be done to better understand how life history and range size interact to shape genetic variation in continuous space, which we leave to future studies.
